---
title: "Regression"
author: "Sahana Rayan"
date: "June 22, 2020"
output: html_document
---

#Loading the dataset
```{r message = FALSE, error=FALSE, warning=FALSE}
library(tidyverse)
tennis_data <- read_csv("http://www.stat.cmu.edu/cmsac/sure/materials/data/regression_projects/tennis_2013_2017_GS.csv")
```

#Cleaning the dataset
When looking at the composition of certain variables, lack of variation was identified.
```{r}
table("Retirement Binary Variable" = tennis_data$Retirement)
table("Winning player's sets won" = tennis_data$w_setswon)
```
 
Some variables were also redundant that required to be removed

```{r}
table("Slam" = tennis_data$slam,
      "Tournament" = tennis_data$tournament)
```


This information helped in creating this cleaned dataframe
```{r}
tennis_data_cleaned <- tennis_data %>% 
  select(-c("w_n_netpt_w", "w_n_netpt", "slam", "player1", "player2", "a1", "a2", "X1", "match_id.y")) %>% 
  rename(Net_points = l_n_netpt,
         Net_points_won = l_n_netpt_w) %>% 
  mutate(w_bp_success = w_n_bp_w/w_n_bp,
         l_bp_success = l_n_bp_w/l_n_bp,
         w_sv_success = w_n_sv_w/w_n_sv,
         l_sv_success = l_n_sv_w/l_n_sv,
         upset_bool = ifelse(winner_rank >loser_rank, TRUE, FALSE)) %>% 
  filter(Retirement == FALSE) %>% 
  filter(w_setswon %in% c(2,3))
```

The __select__ line removed all the redundant variables. The __rename__ function was used to rename certain variables to fit the context of the dataframe. The __mutate__ function was used to add some variables that we though couldbe worth looking at for EDA. These variables are serve success, break point success and a boolean variable that determines whether that particular match was an upset match. The __filter__ was used to filter out data that lacked variation.


#EDA for continuous variables
The correlation matrix was used to look at the linear relationships between certain continuous variables that we were interested in

```{r}
library(ggcorrplot)
tennis_model_data <- tennis_data_cleaned %>%
  dplyr::select(w_n_winners, 
                w_bp_success,
                w_sv_success,
                w_pointswon,
                w_n_ue,
                l_n_ue,
                w_n_sv_w)
tennis_cor_matrix <- cor(tennis_model_data)
ggcorrplot(tennis_cor_matrix)


round_cor_matrix <- 
  round(cor(tennis_model_data), 2)
ggcorrplot(round_cor_matrix, 
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)

```

The losing player's number of unforced errors, winning player's number of winners, winning player's number of serve wins, and winning player's number of unforced errors had a strong correlation with winning player's total points won

*w_pointswon* appeared to be an interesting variable and so the distrubution of this variable looked like this
```{r}
tennis_data_cleaned %>% 
  ggplot() +
  geom_histogram(aes(x = w_pointswon)) +
  xlab("Winning player's points won") +
  theme_bw()
```


The distribution appears to have some outliers and once these outliers are removed, the distribution looks like this:
```{r}
tennis_data_cleaned %>% 
  filter((w_pointswon > 50) & (w_pointswon < 200)) %>%
  ggplot() +
  geom_histogram(aes(x = w_pointswon)) +
  xlab("Winning player's points won") +
  theme_bw()
tennis_data_linear <- tennis_data_cleaned %>% 
  filter((w_pointswon > 50) & (w_pointswon < 200))
```



The scatterplots for the aforementioned variables against *w_pointswon* were plotted
```{r message = FALSE}
library(GGally)
ggpairs(tennis_data_linear,
        columns =
          c("w_pointswon", "w_n_ue", "l_n_ue", "w_n_winners", "w_n_sv_w"),
        mapping = aes(alpha = 0.2)) +
  theme_bw()

```

```{r messsage = FALSE}
stat.lab <- c("Losing player's number of unforced errors", "Winning player's number of serve wins", 
              "Winning player's number of unforced errors", "Winning player's number of winners")
names(stat.lab) <- c("l_n_ue", "w_n_sv_w", "w_n_ue", "w_n_winners")
tennis_data_cleaned %>% 
  pivot_longer(c("w_n_ue", "w_n_winners", "l_n_ue", "w_n_sv_w"),
               names_to = "Stat", 
               values_to = "Values") %>% 
  ggplot(aes(x = Values, y = w_pointswon)) +
  geom_point(alpha = .2) +
  facet_wrap(~Stat, scales = "free_x", labeller = labeller(Stat = stat.lab)) +
  geom_smooth(method = "lm") +
  ylab("Winning player's total points won") +
  theme_bw() +
  labs(
    title = "Different variables vs Winning player's points won scatterplot"
  )
```



#EDA for Discrete Variables

_round_, _tour_, _Tournament_, _Year_, *w_setswon*, are a couple of categorical variables that could potentially affect *w_pointswon*
